Fast time-domain electromagnetic modelling in the frequency domain

### 2020-10-15 ManuscriptTime.tex
Within revision I

Figure \ref{fig:GUI-FFTLog-log} shows exactly the same on a logarithmic scale,
without the interactive GUI-related elements. It can be seen that with the
chosen frequencies the time-domain response starts to divert above about 100\,s
and below 0.3\,s. It also shows the oscillating high-frequency part, which is
hard to interpolate. Figure \ref{fig:GUI-DLF-lin} is the same as
\ref{fig:GUI-FFTLog-lin}, but transformed with the DLF method applying the
81-point sine-cosine filter from \cite{GEO.09.Key}. The same frequencies were
computed as in the FFTLog case and the missing ones interpolated. The error of
the corresponding time-domain response is comparable, so either FFTLog or DLF
can be used.
%
\multiplot{2}{GUI-FFTLog-log,GUI-DLF-lin}{width=\tcwidth}
{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic scale.
(b) Using DLF instead of FFTLog, but computing the same frequencies as for the
FFTLog. The resulting time domain response has comparable accuracy (compare the
error to the error in Figure \ref{fig:GUI-FFTLog-lin}).
}
%
In the above example we used 20 frequencies, but that many would not be
required for the shown offset of 5\,km, a few of the lower and higher
frequencies could be left out. However, with this frequency-selection we can
model a wide range of offsets. This is shown in Figure \ref{fig:multi-offset},
where the above parameters were used to model the responses at offsets
$r=1.5,3,6$, and 12\,km.
%
\plot{multi-offset}{width=\tcwidth}{
  Normalized (a) frequency- and (b) time-domain responses for offsets
  $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
  \ref{fig:GUI-FFTLog-lin}. The coloured circles are the actually computed
  responses, the black dots are the responses which are set to zero or
  interpolated. The black curves are the analytical responses.
}
%

@@

We have shown a method to minimize the required frequencies and their range for
the computation of time-domain CSEM data with a frequency-domain code. This can
significantly reduce the computation time and makes time-domain CSEM modelling
with a frequency-domain code competitive given a robust frequency-domain
solver, a frequency-dependent gridding function that minimizes the required
cells, and a Fourier transform that works on a logarithmic scale. Fast layered
modelling can be used to design the required frequency range, as the Fourier
transform does not know about the dimensionality of the underlying model.
Twenty frequencies or less are usually sufficient for a wide range of offsets.
We believe that our proposed improvements to the previously published methods
make time-domain CSEM modelling with a frequency-domain code even more
attractive than it was already before. We have demonstrated the idea of our
Fourier transform method on CSEM data transformed from the frequency to the
time domain. However, it could equally be applied to the transform from the
time to the frequency domain and to other methods with similar
characteristics.

>>>>

Figure \ref{fig:GUI-FFTLog-log} shows exactly the same on a logarithmic scale,
without the interactive GUI-related elements. It can be seen that with the
chosen frequencies the time-domain response starts to divert above about 100\,s
and below 0.3\,s. It also shows the oscillating high-frequency part, which is
hard to interpolate. Figure \ref{fig:GUI-DLF-lin} is the same as
\ref{fig:GUI-FFTLog-lin}, but transformed with the DLF method applying the
81-point sine-cosine filter from \cite{GEO.09.Key}. The same frequencies were
computed as in the FFTLog case and the missing ones interpolated. The error of
the corresponding time-domain response is comparable, so either FFTLog or DLF
can be used.
%
\multiplot{2}{GUI-FFTLog-log,GUI-DLF-lin}{width=\tcwidth}
{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic scale. (b)
Using DLF instead of FFTLog, but computing the same frequenc\old{ies}\new{y
range} as for the FFTLog. The resulting time domain response has comparable
accuracy (compare the error to the error in Figure \ref{fig:GUI-FFTLog-lin}).
}
%
In the above example we used 20 frequencies, but that many would not be
required for the shown offset of 5\,km, a few of the lower and higher
frequencies could be left out. However, with this frequency-selection we can
model a wide range of offsets. This is shown in Figure \ref{fig:multi-offset},
where the above parameters were used to model the responses at offsets
$r=1.5,3,6$, and 12\,km.
%
\plot{multi-offset}{width=\tcwidth}{
  Normalized (a) frequency- and (b) time-domain responses for offsets
  $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
  \ref{fig:GUI-FFTLog-lin}. The coloured circles are the actually computed
  responses, the black dots are the responses which are set to zero or
  interpolated. The black curves are the analytical responses.
}
%

@@

We have shown a method to minimize the required frequencies and their range for
the computation of time-domain CSEM data with a frequency-domain code. This can
significantly reduce the computation time and makes time-domain CSEM modelling
with a frequency-domain code competitive given a robust frequency-domain
solver, a frequency-dependent gridding function that minimizes the required
cells, and a Fourier transform that works on a logarithmic scale. Fast layered
modelling can be used to design the required frequency range, as the Fourier
transform does not know about the dimensionality of the underlying model.
Twenty frequencies or less are usually sufficient for a wide range of offsets.
\new{The values for lower frequencies can be interpolated using PCHIP knowing
that the imaginary part goes to zero for zero frequency. The values for higher
frequencies can be set to zero, as we can neglect their influence. And values
for frequencies in-between the computed ones are best obtained with a spline
interpolation. The actual transform can be made with either the DLF method or
FFTLog, where the latter one requires usually much fewer frequencies to be
interpolated.} We believe that our proposed improvements to the previously
published methods make time-domain CSEM modelling with a frequency-domain code
even more attractive than it was already before. We have demonstrated the idea
of our Fourier transform method on CSEM data transformed from the frequency to
the time domain. However, it could equally be applied to the transform from the
time to the frequency domain and to other methods with similar
characteristics.
